Master thesis problem statement: Keyword search over heterogeneous sources

### 2012-10-19 doc/thesis/latex/problem_statement.tex
improvements to the text

For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
%
	Even a simple structured query language and entity names over the mediated schema  at first may seem hard to learn%
	\footnote{on the other hand, at CERN, the names in the mediated schema are refer to entities that in the real-world are fairly consistently named (though there may exist slight differences in their naming, whereas DAS is to provide the 'golden' naming convention).}. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
%   
   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.

\subsection{Supporting complex queries}
Currently DAS can only process queries that could be directly mapped to data providers' APIs, but not the queries where results of one API need to be fed into another APIs 
	%\footnote{this is partially because of performance reasons {\color{red}and the fact that DAS is mainly schema-less (the mediated schema is only in terms of entities, their mappings to APIs and API parameters)}}%
	\footnote{except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
. %In such case, the user  is forced to write some code to submit and process the sub-queries himself. 
Further, to be flexible DAS only keeps a list of 'mediated entities' and how they could be retrieved, but do not enforce any predefined schema (i.e. entity fields/attributes) nor exposes it\footnote{%
	Actually the fields of each API  (an entity could be mapped to a number of these) could be inferred 
	from past queries or by predefined 'bootstrap' queries. Currently the field lists may differ slightly depending on the parameters}
before querying for some particular entity(-ies). 
%
This makes it fairly hard to execute complex queries  without a priori knowledge.
User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program does that.
%
To be answered:

	\begin{compactitem}
		\item how should these queries be entered by users?

		\item could we use the current source descriptions that only define API result type and API parameters as entities of mediated schema to select meaningful and sufficiently efficient composition of services?
	\end{compactitem}

@@

DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile (e.g. new records gathered or existing records updated; service descriptions may be changed; new services added), while some subsets of the data could be fairly static (e.g. datasets stored at DBS do not change often). 
%
Further as the total data that is stored at data-providers could be fairly large (order of 1TB per year) and as their main goal is supporting production (data taking from the CMS detector), the providers may impose some constraints (e.g. only certain queries allowed).
%
Therefore, it is important to balance between returning locally cached query results quickly   and between issuing queries to slow data-services to get fresh results.
%
Issues to be addressed includes:
\begin{compactitem}
						\item Low data providers' performance (e.g. on DBS some queries requires joining many very large tables)
	     				\item At the moment all queries are put into one pool and has to wait until some threads are available. If a couple of heavy queries were submitted earlier, even light queries would have to wait long. Explore more advanced Query prioritization 
	     					(e.g. we could have some query cost model based on history or predefined scores per API)

@@

					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no ordering and limit functionality), DAS has to fetch ALL records matching the query instead of only the first page. Although it would be possible to do partial results for data coming from a single data-source (while still loading in background, or the providers (or us) would have to modify their APIs).

					\item Efficient distributed search: as filtering criteria may reside on two or more autonomous sources, given the current APIs much more items than in the final result may need to be fetched from each source. Items that do not meet all selection criteria could be filtered out only afterwards in DAS. To improve the performance, data provider's APIs for the most used queries may need to be redesigned.
                        % \item {\color{red}e.g. displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
	              %\item {\color{red}scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.                        }
\end{compactitem}

{\color{gray}
\subsection{Generic connector accessing relational databases}
Generic connector able to access relational databases with minimal manual integration could be useful for integrating proprietary systems that were not yet integrated and there's no resources available to build APIs. A possible use-case could be the \textit{Prep} database, but this still has to be discussed. 
%
Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account).
There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database that could be used for this.}

% \color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong. The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job. The work is done over here . Feel free to check it out and we can discuss it more.}

>>>>

For an IR system with a wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while not loosing support of fairly complex queries.
%
	Even a simple structured query language plus entity names over the mediated schema  at first may seem hard to learn%
	\footnote{Actually, the names in DAS mediated schema refer to entities fairly consistently named in the real-world.}. Therefore, it shall be useful to guide new users interactively through the process of building the query.
%   
   Supporting non-structured keyword queries is also worth investigation as quite many users reported missing this Google-like search experience.

\subsection{Supporting complex queries}
Currently DAS can only process queries that could be directly mapped to data providers' APIs, but not the queries where results of one API need to be fed into another APIs 
	%\footnote{this is partially because of performance reasons {\color{red}and the fact that DAS is mainly schema-less (the mediated schema is only in terms of entities, their mappings to APIs and API parameters)}}%
	\footnote{Except a couple of manually hard-coded "views" in a virtual \textit{"combined"} data provider}%
. %In such case, the user  is forced to write some code to submit and process the sub-queries himself. 
Further, to be flexible DAS only keeps a list of 'mediated entities' and how they could be retrieved, but do not enforce any predefined schema (i.e. entity fields/attributes) nor exposes it\footnote{%
	Actually the fields of each API  (an entity could be mapped to a number of these) could be inferred 
	from past queries or by predefined 'bootstrap' queries. Currently the field lists may differ slightly depending on the parameters}
before querying for some particular entity(-ies). 
%
This makes it fairly hard to execute complex queries  without a priori knowledge.
User has to know what queries and how have to be combined (depends on the mediated schema and what APIs are available), and has either to combine the results by hand or to write a program doing that.
%
To be researched:
	\begin{compactitem}
		\item how should these queries be entered by users?

		\item could we use current source descriptions that only define API result type and API parameters as entities of mediated schema to select meaningful and sufficiently efficient composition of services?
	\end{compactitem}

@@

DAS is based on \textit{Virtual Integration} where data is left at the sources allowing data to be volatile, e.g. new records gathered or existing ones updated, service's schema changed. Still large amounts of the data are fairly static. 
% %the providers may impose some constraints (e.g. only certain queries allowed).}
As APIs of some data providers' are comparatively slow\footnote{This is partially because of large amounts of data stored at the providers (order of ~1TB per year in total) and as their main goal is supporting data taking from the CMS detector their systems were not optimzed for exploratory queries} it is important to balance between quickly returning items from local cache and getting fresh results. Another approach we are taking is spotting the performance issues at the services (e.g. at DBS some queries requires joining many very large tables, while most of their existing data stays static).
%
Issues to be addressed includes:
\begin{compactitem}
	     				\item At the moment all queries are put into one pool and has to wait until some threads are available. If a couple of heavy queries were submitted earlier, even light queries would have to wait long. Explore more advanced Query prioritization 
	     					(e.g. we could have a query cost model)
						%  based on history or predefined scores per API

@@

					\item Since DAS does aggregation across multiple data providers, given their current APIs (with no \textit{ordering} and \textit{paging} of results), DAS has to fetch ALL records matching the query instead of only the first page\footnote{Although it would be possible to show partial results for data coming from a single data-source (still loading in background, or the providers' APIs would have to be modified)}.		
					\item Efficient distributed search: as filtering criteria may reside on two or more autonomous sources, given the current APIs much more items than in the final result may need to be fetched from each source. Items that do not meet all selection criteria could be filtered out only afterwards in DAS. To improve the performance, data provider's APIs for the most used queries may need to be redesigned.
                        % \item {\color{red}e.g. displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
	              %\item {\color{red}scale testing - if we are storing lots of historical info. MongoDB is not so performant if DB cant fit in memory. One of well known solutions is installing SSD.                        }
\end{compactitem}

\subsection{\color{gray}Generic connector accessing relational databases}

{\color{gray}A generic connector accessing relational databases with minimal human effort could be useful for integrating proprietary systems that were not yet integrated when there's no resources to build the APIs. A possible use-case could be the \textit{Prep} database, but this still has to be discussed. 
%
%Depending on the needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account).
There was a project\footnote{\url{https://github.com/vkuznet/PyQueryBuilder}}   at CERN which tried to generalize DBS-Query Language for any database that could be used as DB side mediator.
}
% \color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong. The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job. The work is done over here . Feel free to check it out and we can discuss it more.}

### 2012-10-19 doc/thesis/latex/proposed_solutions.tex
improvements to the text

\item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
 	\footnote{As DAS uses Boolean IR model we could only rank specific structured queries, but not the results itself.}:
	\begin{itemize}
	\item the algorithm will generate mappings between each keyword to either an entity name, selection attribute name or its value (resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries, users' feedback	 and user's personal profile (e.g. search history, department, etc)
	\item build an inverted index of attribute values with some good full-text search toolkit%
				\footnote{e.g. Xapian (http://xapian.org/) implemented in C++ which that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
	If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess
	\item the system could improve its results from users' feedback (e.g. by HMM, will need to decide)
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.

\item knowing all possible APIs parameter values would be beneficial ({\color{red}but not yet sure if available}%
		\footnote{problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues. Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}) and could be used to improve the keyword search (better mapping from keywords to API inputs) or even the structured search (e.g. query cleaning/term auto-completion)
\begin{itemize}
\item having direct access to DB, one could incrementally index target database\footnote{%
			E.g. an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms. For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }%
	tables mapping it's columns into API parameters (manually or automatically).

%\item may need to exclude very large tables
\end{itemize}
	\end{itemize}
\item (?) JavaScript-based query interpreter to ease query writing: could suggest search attribute and entity attribute names. also taking into account some possible ambiguous namings (auto completion)
\end{itemize}

\subsection{Supporting complex queries}
{\color{red}At the moment this still needs some further analysis/consideration, however it seems even with current data-source definitions could be used for building more complex queries.}

DAS Service definitions contain for each data service (API): mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori). ]
From theoretical (correctness) stand point, this seems mappable  into \textit{Information Manifold's} data source definitions described on Ch.\ref{IM_query_translation} of Literature review.

Other issues to think about contain:
\begin{itemize}
\item how these queries shall be entered by the user and how ambiguity shall be handled especially if queries entered in some simplified (semi-structured) language like DAS Query Language currently is
\item any optimizations/heuristics if multiple possible solutions exist
\item handling complex and heavy queries with large number of results, e.g. lazy loading with pagination (at least at the level of service composition as APIs do not yet support pagination/sorting)
\end{itemize}

@@

\subsubsection*{More intelligent caching}
Some of the data entities instances changes very rarely. For example in DBS system old datasets would never change, while new entities are constantly added (still some of their attributes may change, e.g.  validity of a dataset). 
%
Cached copy may be shown by default while up to date results could be retrieved on user's request. An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
%
On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).

Also consider that:
                			\begin{itemize}
                			%\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			%\item could cache even \textit{expired} data, but warning user
                			\item we could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			% \item Can we automatically figure which fields are static and which are changing?    			 {\color{green} Yes, if we enfore APIs to record modification timestamp} {\color{red} I also meant some automatic algorithm, however that would be more complex and less reliable}
                			\item pre-fetching common queries: determining manually and/or automatically
                			\end{itemize}

%%%%%  -------------------------------------------
\subsubsection*{Data providers performance issues}
In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for often used queries.

Use either \textit{materialized refresh fast views with query rewriting} (Oracle; completely transparent for proprietary apps)\cite{Oracle11}
 or some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) to improve performance of heavy queries containing joins and/or aggregations.

@@

Having a materialized view with all these tables joined together would allow answering such queries much quicker. Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with only expense of just couple of times of space.

@@

\subsubsection*{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side [pure SQL is possible for MySQL, to check for Oracle]- take a query and bit-vector as parameter) , lazy pagination (and order required for aggregation) - this is not yet supported by any of the data-service APIs

%\newline
% integration at DAS level. (at source DB could be more performant a little)

>>>>

\item JavaScript-based query interpreter with auto completion to ease query writing: could suggest searchable entities and maybe even entity attribute names.

\item Given a keyword query, suggest the best matching structured queries (as a ranked list)%
 	\footnote{As DAS uses Boolean IR model we could only rank specific structured queries, but not the results itself.}:
	\begin{itemize}
	\item the algorithm will generate mappings between each keyword to either an entity name, selection attribute name or its value (resembling 'Query forms' and 'Keyword cleaning' approaches)
	\item ranking could be based on: how closely keywords are matching (some or all) required API's input parameters, popularity of certain queries, users' feedback	 and user's personal profile (e.g. search history, department, etc)
	\item build an inverted index of attribute values with some good full-text search toolkit%
				\footnote{e.g. Xapian (http://xapian.org/) that is known for its flexibility, or Sphinx (http://sphinxsearch.com/)}. 
	If no match found (e.g. new entry not yet in our cache) pattern matching could be used to make a best guess
	\item the system could improve its results from users' feedback through machine learning (e.g. HMM)
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{%
		Matching a wildcard query into a schema field will be harder. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants to allow more flexible queries, at least for \textit{datasets}}.

\item knowing all possible APIs' parameter values, in addition to historical values seen so far, would improve quality of the keyword search (better mapping from keywords to these values) or even the structured search e.g. through query cleaning \& auto-completion. {\color{red}Not yet known if available}%
		\footnote{problem: the owners of services may not want to provide direct access to DB because of no trust, security or performance issues. Otherwise, specific APIs could be developed to return possible parameter values but this is quite an overhead}.
\begin{itemize}
\item having direct access to DB, one could incrementally index target database\footnote{%
			E.g. an inverted index of values in Oracle/MySQL DB tables can be built to keep the list of possible terms. For instance, Sphinx full text search engine can access DBs directly and also supports incremental indexing but it needs a bit of manual configuration {http://sphinxsearch.com/docs/current.html\#delta-updates} }
	tables semi-automatically mapping it's columns into API parameters.

%\item may need to exclude very large tables
\end{itemize}
	\end{itemize}
\end{itemize}

\subsection{Supporting complex queries}
{\color{red}At the moment this still needs some further analysis/consideration, however it seems even the existing data-source definitions could be used for building more complex queries.}

DAS Service definitions contain: for each data service (API), mapping of its parameters and result type into entities of mediated schema (but NOT ALL their fields are defined a priori). Even if DAS definitions do not describe every attribute of API result, as all input parameters are attributed to a mediated schema entity, these definitions seem mappable  into \textit{Information Manifold's} data source definitions described on Sec.\ref{IM_query_translation} of Literature review.

Other issues to consider:
\begin{itemize}
\item how these queries shall be entered by the user and how ambiguity shall be handled especially if queries entered in some simplified (semi-structured) language like DAS Query Language currently is
\item any optimizations/heuristics if multiple possible solutions exist
\item handling complex and heavy queries with large number of results, e.g. lazy loading with pagination (at least at the level of service composition as APIs do not yet support pagination/sorting)
\end{itemize}

@@

\subsubsection*{More intelligent caching}
Some of the data instances changes very rarely. For example in DBS system old datasets would never change, while new entities are constantly added (some of their attributes may change, but only rarely, e.g.  validity of a dataset). 
%
Cached copy may be shown by default while up to date results could be retrieved on user's request. An automatic change rate prediction could be useful to efficiently balance between caching and retrieving results.
%
On the other hand some result items may not make sense to be cached, as there may be too many of valid input parameter combinations (e.g. now there are 900M of \textit{(run, file)} combinations).

Other way to boost performance is pre-fetching most common queries from time to time (determined manually and/or automatically). Also it's possible have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used

%%%%%  -------------------------------------------
\subsubsection*{Data providers' performance issues}
In the cases when new records are coming but the existing ones are not changing much, continuous view maintenance that computes only differences from earlier results could be a good solution to improve the performance especially for popular and heavy queries containing joins and/or aggregations.

@@

Having a materialized view with all these tables joined together would allow answering such queries much quicker. Given low change rates (in comparison to data already present), maintaining the view should be also comparatively cheap with the only expense of just couple of times of storage space (storage is bound by the size of the largest table anyway).

For Oracle \textit{materialized refresh fast views with query rewriting} would be completely transparent and would not need any changes to proprietary systems, but it has a couple of limitations on the queries\cite{Oracle11}. Otherwise some other continuous view maintenance tool (e.g. DBToaster\footnote{http://www.dbtoaster.org that is being developed at EPFL}) could be used, however this wouldn't be as transparent.

@@

\subsubsection*{Integrating distributed information efficiently}
Bloom-join (which could be quite transparent and implemented even on DB side (efficient pure SQL is doable for MySQL, and probably for Oracle too) quite almost transparently (a helper function could take the initial query and bloom-filter's bit-vector as parameters).

### 2012-10-19 doc/thesis/latex/report_sept.tex
improvements to the text

\addtolength{\topmargin}{-10mm}
\addtolength{\topskip}{-10mm}

@@

\begin{document}

@@

\section{Introduction}
At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
\footnote{For instance, at CERN,  due to many reasons (e.g. research and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}. Further high turnover makes it harder extending these systems}. As data stored on one system may be related to data residing on other systems%
	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the worldwide grid storage}%
, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.

Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. A data integration system based on simple structured queries is already in place. Various improvements including support for less restricted keyword queries, improvements to system's usability and performance still have to be researched.

@@

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find (all) \textit{datasets} related some specific physics phenomena\footnote{In case of dataset this data is present in filename or run} together with conditions describing how this data was recorded by detector or simulated which are present in separate autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} 
         	  \begin{small}(filtering attributes from separate services)\end{small}
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

@@

The Data Aggregation System (DAS)\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema ('DAS keys') into web-service methods and their arguments. Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use. DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]) operators. The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:

{\small 
\begin{verbatim}
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{verbatim}
}

The query above would return average and median datasets sizes  of ones containing  \textit{RelVal} in their name having more than 1000 events.

Queries could be run either from web browser or through  command line interface where the results could  be fed into another application (e.g. program doing physics analysis or automatic software release validation).

% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}

\input{problem_statement}

>>>>

%\addtolength{\topmargin}{-10mm}
%\addtolength{\topskip}{-10mm}

@@

\begin{document}
%\addtolength{\bottommargin}{-10mm}
\addtolength{\textheight}{10mm}

@@

\section{Introduction}
At large scientific collaborations like the CMS Experiment at CERN's LHC that includes more than 3000 collaborators data usually resides on a fair number of autonomous and heterogeneous proprietary systems each serving it's own purpose
\footnote{For instance, at CERN,  due to many reasons (e.g. research-orientedness and need of freedom, politics of institutes involved) software projects usually evolve in independent fashion resulting in fair number of proprietary systems\cite{Koch00CERN}. Further high turnover makes it harder to extend these systems}. As data stored on one system may be related to data residing on others%
	\footnote{For example, datasets containing physics events are registered at DBS, while the physical location of files is tracked by Phedex which also takes care of their transfers within the grid storage worldwide}%
, users are in need of a centralized and easy-to-use solution for locating and combining data from all these multiple services.

Using a highly structured language like SQL is problematic because users need to know not only the language but also where to find the information and also lots of technical details like schema. An enterprise information integration system based on simple structured queries is already in place. Various improvements including: adding support for more complex queries and for less restricted keyword search, improvements to system's usability and performance still have to be researched.

@@

\section{Case study: the CMS Experiment at CERN}
Users' information need may vary greatly depending on their role, however most of the time they are interested in locating a \textit{full set} of entities matching some selection criteria, e.g.:
%\footnote{in contrary to exploratory approach where  a subset of best matches is enough}
   \begin{itemize}
  		\item find all \textit{files} from \textit{dataset(s)} matching wild-card query each containing some of the 'interesting' \textit{runs} from a list provided (Release validation teams)
         \item find all \textit{datasets} matching some pattern stored at a given \textit{site} 
         	  \begin{small}(filtering attributes from separate services)\end{small}
         \item find (all) \textit{datasets} related to some specific physics phenomena\footnote{In case of dataset, this data is present in the \textit{filename} or in the \textit{run} entity} together with conditions describing how this data was recorded by detector or simulated which are present in another autonomous system than the datasets (Physicists)
         % TODO:is this a good example? conditions is a separate query now...
   \end{itemize}                		
% In such dynamic environment it is preferred to choose local-as-view approach for information integration instead of creating a global schemata\cite{Koch00CERN}.

@@

The \textit{CMS Data Aggregation System (DAS)}\cite{Kuznetsov2010, Kuznetsov2011} was created which allows integrated access to a number of proprietary data-sources by processing user's queries on demand - it determines  data-sources are able to answer\footnote{This is done by a mapping between flat mediated schema entities ('DAS keys') into web-service methods and their arguments. Then system queries all services that could provide a result of expected type with given parameters}, queries them, merges the results and caches them for subsequent use. DAS uses \textit{Boolean retrieval model} as users are often interested in retrieving ALL the items matching their query.

Currently the queries specify what entity the user is interested in (dataset, file, etc) and provide selection criteria (attribute=value, name BETWEEN [v1, v2]). The combined query results could be later 'piped' for further filtering, sorting or aggregation (min, max, avg, sum, count, median), e.g.:

{\small 
\begin{Verbatim}[commandchars=\\\{\}]
\footnotesize# returns average and median dataset sizes of ones containing \textit{RelVal} in their name having more than 1000 events.
dataset=*RelVal* | grep dataset.nevents >1000 | avg(dataset.size), median(dataset.size)
\end{Verbatim}
}

Queries are executed either from web browser or through a command line interface where the results could be fed into another application (e.g. program doing physics analysis or automatic software release validation).
% \textbf{\color{red}Examples of more general uses: Digital libraries, Stock markets, [ Weather, Flight costs]}
\input{problem_statement}
